---
title: "daca_regressions_v2"
output: html_document
date: "2023-04-15"
editor_options: 
  chunk_output_type: console
---

- DACA eligibility criteria:
  - Under the age of 31 as of June 15, 2012 (Born on or after 6/16/1981);
  - Came to the United States before their 16th birthday;
  - Have continuously resided in the United States since June 15, 2007, up to the present time;
  - Were physically present in the United States on June 15, 2012, and at the time of making their request for consideration of deferred action with USCIS;
  - Had no lawful status on June 15, 2012;
  - Are currently in school, have graduated or obtained a certificate of completion from high school, have obtained a general education development (GED) certificate, or are an honorably discharged veteran of the Coast Guard or Armed Forces of the United States; and
  - Have not been convicted of a felony, significant misdemeanor, or three or more other misdemeanors, and do not otherwise pose a threat to national security or public safety.
  
  # Importing & Setup
```{r message=FALSE, warning=FALSE}
# setwd
#setwd("~")

# import library
library(tidyverse)
library(fixest)
library(modelsummary)

# import datasets
df <- read_csv("new_data/clean_non_cit_cali_2005_2021.csv")
df <- df |> 
  rename(year = survey_year)

# mexicans only
mex_df <- df |> 
  filter(hisp == 2,
         year >= 2008) |> 
  rename(age = agep) |> 
  mutate(hs_only = ifelse(schl == 16, 1, 0))
```


# Sample 3: All non-citizens ages 18-40
```{r}
sample_3 <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18)
```


### Working (dummy variable)
```{r}
# creating DiD
did_sample_3_working <- feols(working ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Unemployed
```{r}
# creating DiD
did_sample_3_unemployed <- feols(unemployed ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + years_of_education + male + married | year +
                     puma, data = sample_3)
```


### Income: pernp
```{r}
# creating DiD
did_sample_3_pernp <- feols(pernp ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age^2 + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Income: log(pernp + 1)
```{r}
did_sample_3_pernp_log <- feols(log(pernp + 1) ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age^2 + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Hours per week: wkhp
```{r}
# creating DiD
did_sample_3_wkhp <- feols(wkhp ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Self-employed
```{r}
# creating DiD
did_sample_3_self_employed <- feols(self_employed ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### high school or equivalent
```{r}
# creating DiD
did_sample_3_hs_or_equiv <- feols(high_school_and_equiv ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + male + married +
                   unemployed | year + puma, data = sample_3)
```


### regular high school degree only
```{r}
did_sample_3_hs_only <- feols(hs_only ~ interaction_daca_eligible_post_daca + daca_eligible +
                                post_daca + age + male + married +
                                unemployed | year + puma, data = sample_3)
```


### GED
```{r}
# creating DiD
did_sample_3_ged <- feols(ged ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca + age + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Attended college
```{r}
# creating DiD
did_sample_3_attend_college <- feols(attended_college  ~ interaction_daca_eligible_post_daca +
                                       daca_eligible + post_daca + age + male + 
                                       married + age + unemployed | year
                                     + puma, data = sample_3)
```


```{r}
# getting coefficient names to rename
coef_names <- c("daca eligible * post daca", "daca eligible")

# create panel
models <- list(
    "Worked" = did_sample_3_working,
    "Unemployed" = did_sample_3_unemployed,
    "Income" = did_sample_3_pernp,
    "Log(Income + 1)" = did_sample_3_pernp_log,
    "Hours Worked" = did_sample_3_wkhp,
    "Self Employed" = did_sample_3_self_employed,
    "High School or Equivalent" = did_sample_3_hs_or_equiv,
    "High School Only" = did_sample_3_hs_only,
    "GED" = did_sample_3_ged,
    "Attended College" = did_sample_3_attend_college
)

modelsummary(models)

# creating reg output panels
### separate education and labor outcomes
modelsummary(
  models,
  coef_omit = c(-1, -2),
  coef_rename = coef_names,
  estimate = "{estimate}{stars}",
  gof_map = c("nobs", "r.squared"),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title = "Non-Citizen Mexicans Ages 18-40",
  notes = "This table reports the estimate from the DiD equation with all controls and fixed effects. ***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level."
)
```

```{r}
# getting coefficient names to rename
coef_names <- c("daca eligible * post daca", "daca eligible")

# create panel
labor_models <- list(
    "Worked" = did_sample_3_working,
    "Unemployed" = did_sample_3_unemployed,
    "Income" = did_sample_3_pernp,
    "Hours Worked" = did_sample_3_wkhp,
    "Self Employed" = did_sample_3_self_employed
)

education_models <- list(
    #"High School or Equivalent" = did_sample_3_hs_or_equiv,
    #"High School Only" = did_sample_3_hs_only,
    "GED" = did_sample_3_ged,
    "Attended College" = did_sample_3_attend_college
)

# creating reg output panels
### separate labor outcomes
modelsummary(
  labor_models,
  coef_omit = c(-1, -2),
  coef_rename = coef_names,
  estimate = "{estimate}{stars}",
  gof_map = c("nobs", "r.squared", "vcov.type"),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title = "Non-Citizen Mexicans Ages 18-40: Labor Outcomes",
  notes = "This table reports the estimate from the DiD equation with all controls and fixed effects. ***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level."
)

### educational outcomes
modelsummary(
  education_models,
  coef_omit = c(-1, -2),
  coef_rename = coef_names,
  estimate = "{estimate}{stars}",
  gof_map = c("nobs", "r.squared", "vcov.type"),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title = "Non-Citizen Mexicans Ages 18-40: Educational Outcomes",
  notes = "This table reports the estimate from the DiD equation with all controls and fixed effects. ***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level."
)
```


###############################################################################
# Linear Trend Lines ##########################################################
###############################################################################

### Working (dummy variable)
```{r}
# creating DiD
est_did_pernp <- feols(pernp ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed
                 | year + puma + age_of_entry, data = sample_3)

# plotting linear time trend
iplot(est_did_pernp,
      main = "Effect on Income",
      xlab = "Year")

# creating DiD
est_did_pernp_log <- feols(log(pernp + 1) ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_pernp_log,
      main = "Effect on Log(income + 1)",
      xlab = "Year")


# creating DiD
est_did_wkhp <- feols(wkhp ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed
                 | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_wkhp,
      main = "Effect on Hours Worked per Week",
      xlab = "Year")

# creating DiD
est_did_attended_college <- feols(attended_college ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_attended_college,
      main = "Effect on College Attendance",
      xlab = "Year")

# creating DiD
est_did_working <- feols(working ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_working,
      main = "Effect on Working",
      xlab = "Year")

# creating DiD
est_did_unemployed <- feols(unemployed ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_unemployed,
      main = "Effect on Unemployment",
      xlab = "Year")

# creating DiD
est_did_self_employed <- feols(self_employed ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + age^2 + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_self_employed,
      main = "Effect on Self Employment",
      xlab = "Year")

# creating DiD
est_did_hs <- feols(high_school_and_equiv ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_hs,
      main = "Effect on High School or Equivalent",
      xlab = "Year")

# creating DiD
est_did_ged <- feols(ged ~ i(factor_var = year, var = interaction_daca_eligible_post_daca) + daca_eligible + post_daca + years_of_education + male + married + unemployed | year + puma, data = sample_3)

# plotting linear time trend
iplot(est_did_ged,
      main = "Effect on GED",
      xlab = "Year")
```

# Regression without control
### Working (dummy variable)
```{r}
# creating DiD
did_sample_3_working_no_cont <- lm(working ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Unemployed
```{r}
# creating DiD
did_sample_3_unemployed_no_cont <- lm(unemployed ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Income: pernp
```{r}
# creating DiD
did_sample_3_pernp_no_cont <- lm(pernp ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Income: log(pernp + 1)
```{r}
did_sample_3_pernp_log_no_cont <- lm(log(pernp + 1) ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Hours per week: wkhp
```{r}
# creating DiD
did_sample_3_wkhp_no_cont <- lm(wkhp ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Self-employed
```{r}
# creating DiD
did_sample_3_self_employed_no_cont <- lm(self_employed ~ interaction_daca_eligible_post_daca +
                                              daca_eligible + post_daca, data = sample_3)
```


### high school or equivalent
```{r}
# creating DiD
did_sample_3_hs_or_equiv_no_cont <- lm(high_school_and_equiv ~ interaction_daca_eligible_post_daca + daca_eligible + post_daca, data = sample_3)
```


### regular high school degree only
```{r}
did_sample_3_hs_only_no_cont <- lm(hs_only ~ interaction_daca_eligible_post_daca + daca_eligible +
                                post_daca, data = sample_3)
```


### GED
```{r}
# creating DiD
did_sample_3_ged_no_cont <- lm(ged ~ interaction_daca_eligible_post_daca + daca_eligible +
                   post_daca, data = sample_3)
```


### Attended college
```{r}
# creating DiD
did_sample_3_attend_college_no_cont <- lm(attended_college  ~ interaction_daca_eligible_post_daca +
                                       daca_eligible + post_daca, data = sample_3)
```


```{r}
# getting coefficient names to rename
coef_names <- c("daca eligible * post daca", "daca eligible")

# create panel
labor_models_no_control <- list(
    "Worked" = did_sample_3_working_no_cont,
    "Unemployed" = did_sample_3_unemployed_no_cont,
    "Income" = did_sample_3_pernp_no_cont,
    "Hours Worked" = did_sample_3_wkhp_no_cont,
    "Self Employed" = did_sample_3_self_employed_no_cont
)

education_models_no_control <- list(
    #"High School or Equivalent" = did_sample_3_hs_or_equiv_no_cont,
    #"High School Only" = did_sample_3_hs_only_no_cont,
    "GED" = did_sample_3_ged_no_cont,
    "Attended College" = did_sample_3_attend_college_no_cont
)

# creating reg output panels
### separate labor outcomes
modelsummary(
  labor_models_no_control,
  coef_omit = c(-1, -2),
  coef_rename = coef_names,
  estimate = "{estimate}{stars}",
  gof_map = c("nobs", "r.squared"),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title = "Non-Citizen Mexicans Ages 18-40: Labor Outcomes",
  notes = "This table reports the estimate from the DiD equation without controls and fixed effects. ***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level."
)

### educational outcomes
modelsummary(
  education_models_no_control,
  coef_omit = c(-1, -2),
  coef_rename = coef_names,
  estimate = "{estimate}{stars}",
  gof_map = c("nobs", "r.squared"),
  stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
  title = "Non-Citizen Mexicans Ages 18-40: Educational Outcomes",
  notes = "This table reports the estimate from the DiD equation without controls and fixed effects. ***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level."
)
```

# Pre-trends table

### Working (dummy variable)
```{r}
# creating DiD
did_sample_3_working_check <- feols(working ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Unemployed
```{r}
# creating DiD
did_sample_3_unemployed_check <- feols(unemployed ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + years_of_education + male + married | year +
                     puma, data = sample_3)
```


### Income: pernp
```{r}
# creating DiD
did_sample_3_pernp_check <- feols(pernp ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age^2 + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Income: log(pernp + 1)
```{r}
did_sample_3_pernp_log_check <- feols(log(pernp + 1) ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age^2 + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Hours per week: wkhp
```{r}
# creating DiD
did_sample_3_wkhp_check <- feols(wkhp ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Self-employed
```{r}
# creating DiD
did_sample_3_self_employed_check <- feols(self_employed ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + years_of_education + male + married +
                   unemployed | year + puma, data = sample_3)
```


### high school or equivalent
```{r}
# creating DiD
did_sample_3_hs_or_equiv_check <- feols(high_school_and_equiv ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + male + married +
                   unemployed | year + puma, data = sample_3)
```


### regular high school degree only
```{r}
did_sample_3_hs_only_check <- feols(hs_only ~ i(year, daca_eligible, 2012) + daca_eligible +
                                post_daca + age + male + married +
                                unemployed | year + puma, data = sample_3)
```


### GED
```{r}
# creating DiD
did_sample_3_ged_check <- feols(ged ~ i(year, daca_eligible, 2012) + daca_eligible +
                   post_daca + age + male + married +
                   unemployed | year + puma, data = sample_3)
```


### Attended college
```{r}
# creating DiD
did_sample_3_attend_college_check <- feols(attended_college  ~ i(year, daca_eligible, 2012) +
                                       daca_eligible + post_daca + age + male + 
                                       married + age + unemployed | year
                                     + puma, data = sample_3)
iplot(did_sample_3_wkhp_check)
```


```{r}
# getting coefficient names to rename
coef_names <- c("Eligible * 2008", "Eligible * 2009", "Eligible * 2010", "Eligible * 2011",
                "Eligible * 2013", "Eligible * 2014", "Eligible * 2015", "Eligible * 2016",
                "Eligible * 2017", "Eligible * 2018", "Eligible * 2019", "Eligible * 2021")

# create panel
models <- list(
    "Worked" = did_sample_3_working_check,
    "Unemployed" = did_sample_3_unemployed_check,
    "Income" = did_sample_3_pernp_check,
    #"Log(Income + 1)" = did_sample_3_pernp_log_check,
    "Hours Worked" = did_sample_3_wkhp_check,
    "Self Employed" = did_sample_3_self_employed_check,
    #"High School or Equivalent" = did_sample_3_hs_or_equiv_check,
    #"High School Only" = did_sample_3_hs_only_check,
    "GED" = did_sample_3_ged_check,
    "Attended College" = did_sample_3_attend_college_check
)

#### solid starting point for pre-trends table here
###### remove everything except the years
modelsummary(models,
             coef_omit = c(-1:-12),
             coef_rename = coef_names,
             estimate = "{estimate}{stars}",
             gof_map = c("nobs", "r.squared", "vcov.type"),
             stars = c("*" = 0.1, "**" = 0.05, "***" = 0.01),
             title = "Pre Trends",
             notes = "This table reports the estimate with eligible interacted with each year. The 2012 interaction is the omitted interaction. The DiD equation contains all controls and fixed effects. \n***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level.")


# plots
iplot(did_sample_3_working_check,
      main = "Pre Trends: Worked")
iplot(did_sample_3_unemployed_check,
      main = "Pre Trends: Unemployed")
iplot(did_sample_3_pernp_check,
      main = "Pre Trends: Income")
iplot(did_sample_3_wkhp_check,
      main = "Pre Trends: Hours Worked Per Week")
iplot(did_sample_3_self_employed_check,
      main = "Pre Trends: Self Employed")
iplot(did_sample_3_ged_check,
      main = "Pre Trends: GED")
iplot(did_sample_3_attend_college_check,
      main = "Pre Trends: Attended College")

```


