---
title: "eda"
author: "Oscar Alonso"
date: "2023-03-24"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

- Outcome Variables of Interest
  - Income: WAGP, PERNP
  - Hours worked: WKHP, COW
  - Education: SCHL
  
- DACA eligibility criteria:
  - Under the age of 31 as of June 15, 2012 (Born on or after 6/16/1981);
  - Came to the United States before their 16th birthday;
  - Have continuously resided in the United States since June 15, 2007, up to the present time;
  - Were physically present in the United States on June 15, 2012, and at the time of making their request for consideration of deferred action with USCIS;
  - Had no lawful status on June 15, 2012;
  - Are currently in school, have graduated or obtained a certificate of completion from high school, have obtained a general education development (GED) certificate, or are an honorably discharged veteran of the Coast Guard or Armed Forces of the United States; and
  - Have not been convicted of a felony, significant misdemeanor, or three or more other misdemeanors, and do not otherwise pose a threat to national security or public safety.
  
- Variables to control for
  - Pre-existing differences in education levels, income, and other socioeconomic factors between DACA-eligible and ineligible individuals prior to DACA implementation.
  
  - Changes in immigration policy and enforcement during the DACA implementation period, which could affect outcomes for both DACA-eligible and ineligible individuals.
  
  - The timing of the survey and potential changes in economic or social conditions during the survey period that could affect outcomes.
  
  - Differences in the composition of the Mexican population before and after DACA implementation, such as changes in age distribution or geographic location.
  
  - The potential for selection bias if individuals who apply for DACA differ systematically from those who do not.
  
  - Differences in the length of time spent in the U.S. between DACA-eligible and ineligible individuals, which could affect outcomes.
  
  - The potential for unmeasured variables that may affect outcomes and are correlated with DACA eligibility.

- regression table output: https://www.jakeruss.com/cheatsheets/stargazer/
- another example: look at qje https://www.jakeruss.com/cheatsheets/stargazer/

```{r}
# libraries
library(tidyverse)
library(tmap)
library(tigris)
options(tigris_use_cache = TRUE)
library(modelsummary)

# setwd
setwd("~/Desktop/thesis/notebooks/")

# importing data
df <- read_csv("new_data/clean_non_cit_cali_2005_2021.csv")
```

```{r}
# creating mexican only df
mex_df <- df |> 
  filter(hisp == 2) |> 
  rename(age = agep)
```

# EDA
```{r}
# where do daca eligible people come from?
df |> 
  filter(daca_eligible == 1,
         hisp_label != "not spanish/latino/hispanic",
         hisp_label != "All Other Spanish/Hispanic/Latino") |> 
  group_by(hisp_label) |> 
  summarise(n = n()) |> 
  arrange(-n) |> 
  mutate(pct_share = n / sum(n)) |> 
  head(5) |> 
  ggplot(aes(x = fct_reorder(hisp_label, n), y = pct_share)) +
  geom_col() +
  scale_y_continuous(n.breaks = 10, labels = scales::percent_format()) +
  coord_flip() +
  labs(title = "Total Daca Eligible Hispanics in California by Country of Birth",
       caption = "Source: American Community Survey") +
  xlab(NULL) +
  ylab("Total") +
  ggthemes::theme_clean() +
  theme(plot.margin = unit(c(1, 2, 1, 1), "cm"))
```

# Given that most daca eligible people come from Mexico, going to be using only Mexicans

# creating map using pumas
```{r}
### create california map
#### getting outline of CA pumas
ca_pumas <- pumas("CA", year = 2018)

#### plotting the outline of CA
# plot(ca_pumas$geometry)

### getting mex_df for map
ca_data_for_map <- mex_df |> 
  group_by(daca_eligible, puma) |> 
  summarise(total_daca_eligible = sum(daca_eligible * pwgtp)) |> 
  filter(!is.na(total_daca_eligible))

# joining data
ca_pumas <- ca_pumas |> 
  mutate(puma = as.double(PUMACE10))

joined_pumas <- ca_pumas %>%
  left_join(ca_data_for_map, by = "puma")

# creating map
tm_shape(joined_pumas) + 
  tm_polygons(col = "total_daca_eligible",
              palette = "Reds",
              border.alpha = 0.1,
              title = "Total Mexican Daca Eligible People") + 
  tm_layout(legend.outside = TRUE,
            legend.outside.position = "right")

#### top 10 PUMA areas where DACA people are located ####
joined_pumas |> 
  as.data.frame() |> 
  select(NAMELSAD10, total_daca_eligible) |> 
  arrange(-total_daca_eligible) |> 
  head(10) |> 
  mutate(pct_total_daca_share = total_daca_eligible / sum(total_daca_eligible)) |> 
  ggplot(aes(x = fct_reorder(NAMELSAD10, total_daca_eligible),
             y = pct_total_daca_share)) +
  geom_col() + 
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(title = "Total Mexican DACA Eligible People by California PUMA",
       subtitle = "Data as of 2021") +
  xlab(NULL) +
  ylab("Count") +
  ggthemes::theme_clean()

### top 10 counties with the largest % of DACA eligible people ###
county_plot <- joined_pumas |> 
  mutate(county = str_extract(str_to_lower(NAMELSAD10), ".*(?=county)")) |>
  select(county, daca_eligible, total_daca_eligible) |> 
  data_frame() |>
  group_by(county) |> 
  summarise(total = sum(total_daca_eligible)) |> 
  mutate(prop = total / sum(total)) |> 
  arrange(-prop) |> 
  head(10) |> 
  ggplot(aes(x = fct_reorder(county, prop), y = prop)) +
  geom_col() +
  geom_text(aes(label = paste0(scales::percent(prop), "\n")), hjust = 1, 
            vjust = 0.8, color = "white", size = 3) +
  scale_y_continuous(n.breaks = 15, labels = scales::percent_format()) +
  xlab("County") +
  ylab("Proportion") +
  labs(title = "Percent of DACA Eligible People by County",
       subtitle = "Most are located in southern California") +
  coord_flip() +
  ggthemes::theme_clean()
```

# creating tables and plots
```{r}
### Creating data summary df ###
mex_df_summary <- mex_df |> 
  mutate(daca_eligible_label = ifelse(daca_eligible == 1, "Daca Eligible", "Ineligible")) |>
  select(daca_eligible_label, age, age_of_entry, yoep, pernp,
         wkhp, years_of_education, years_living_in_us) |> 
  rename(
    'age of entry' = age_of_entry,
    'year of entry' = yoep,
    income = pernp,
    'hours worked per week' = wkhp,
    'years of education' = years_of_education,
    'years living in US' = years_living_in_us
  )

# creating data summary table
datasummary_balance(
  ~daca_eligible_label,
  data = mex_df_summary,
  fmt = 0,
  title = "Non-citizen Mexican DACA Eligible vs. Ineligible in California",
  stars = c('*' = .1, '**' = .05, '***' = 0.01),
  notes = "***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level.",
  dinm_statistic = "p.value"
)


# creating income table
income_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(mean_income = mean(pernp, na.rm = T),
            se_income = sd(pernp, na.rm = T) / sqrt(n())) |> 
  pivot_wider(names_from = daca_eligible, values_from = c(mean_income, se_income)) |> 
  # rename(non_eligible = '0',
  #        eligible = '1',
  #        non_eligible_se = '0',
  #        eligible_se = '1') |> 
  mutate(eligible_minus_non = mean_income_1 - mean_income_0,
         eligible_minus_non_se = sqrt(se_income_1^2 + se_income_0^2)) |> 
  select(survey_year, eligible_minus_non, eligible_minus_non_se)

 # plotting income table
  income_table |> 
    filter(survey_year >= 2008) |> 
    ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - eligible_minus_non_se,
                    ymax = eligible_minus_non + eligible_minus_non_se)) +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(labels = scales::dollar_format(), n.breaks = 10) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Income") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))
  
# creating income table
income_log_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  mutate(income_log = log(pernp + 1)) |> 
  summarise(mean_income = mean(income_log, na.rm = T)) |> 
  pivot_wider(names_from = daca_eligible, values_from = mean_income) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

 # plotting income table
income_log_table |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
  geom_line(size = 1) +
  geom_point() +
  scale_x_continuous(n.breaks = 20) +
  scale_y_continuous(labels = scales::dollar_format(), n.breaks = 10) +
  geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
            fill = "gray40", alpha = 0.05) +
  labs(title = "Log(Income + 1)") +
  xlab("Year") +
  ylab("Difference Between Eligible and Ineligible") +
  ggthemes::theme_clean()

# creating hours worked table
wkhp_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(mean_wkhp = mean(wkhp, na.rm = T),
            se = sd(wkhp, na.rm = T) / sqrt(n())) |> 
  pivot_wider(names_from = daca_eligible, values_from = c(mean_wkhp, se)) |> 
  # rename(non_eligible = '0',
  #        eligible = '1') |> 
  mutate(eligible_minus_non = mean_wkhp_1 - mean_wkhp_0,
         se_wkhp = sqrt(se_1^2 + se_0^2)) |> 
  select(survey_year, eligible_minus_non, se_wkhp)

# plotting hours worked table
wkhp_table |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se_wkhp,
                    ymax = eligible_minus_non + se_wkhp)) +
    scale_x_continuous(n.breaks = 20) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Hours Worked Per Week") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))

# creating fraction in college
fraction_college_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_atten_college = sum(attended_college == 1) / n(),
            se = sd(attended_college, na.rm = T) / sqrt(n())) |> 
  pivot_wider(names_from = daca_eligible, values_from = c(fraction_atten_college, se)) |> 
  # rename(non_eligible = '0',
  #        eligible = '1') |> 
  mutate(eligible_minus_non = fraction_atten_college_1 - fraction_atten_college_0,
         se = sqrt(se_1^2 + se_0^2)) |> 
  select(survey_year, eligible_minus_non, se)

get_table <- function(df, var) {
  {{ df }} |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction = sum({{ var }} == 1) / n(),
            se = sd({{ var }}, na.rm = T) / sqrt(n())) |> 
  pivot_wider(names_from = daca_eligible, values_from = c(fraction, se)) |> 
  # rename(non_eligible = '0',
  #        eligible = '1') |> 
  mutate(eligible_minus_non = fraction_1 - fraction_0,
         se = sqrt(se_1^2 + se_0^2)) |> 
  select(survey_year, eligible_minus_non, se)
}

# creating fraction working
fraction_working_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_working = sum(working == 1) / n()) |> 
  pivot_wider(names_from = daca_eligible, values_from = fraction_working) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

get_table(mex_df, working)

# plotting fraction working table
get_table(mex_df, working) |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se,
                  ymax = eligible_minus_non + se)) +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(-0.1, 0.1)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Fraction Worked") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))

# creating fraction unemployed
fraction_unemployed_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_unemployed = sum(unemployed == 1) / n()) |> 
  pivot_wider(names_from = daca_eligible, values_from = fraction_unemployed) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

get_table(mex_df, unemployed)

# plotting fraction unemployed table
get_table(mex_df, unemployed) |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se,
                      ymax = eligible_minus_non + se)) +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(-0.08, 0.08)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
  geom_hline(yintercept = 0, color = "red") +
    labs(title = "Fraction Unemployed") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))

# creating fraction self employed
fraction_self_employed_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_self_employed = sum(self_employed == 1) / n()) |> 
  pivot_wider(names_from = daca_eligible, values_from = fraction_self_employed) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

get_table(mex_df, self_employed)

# plotting fraction self employed
get_table(mex_df, self_employed) |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se,
                      ymax = eligible_minus_non + se)) +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(-0.1, 0.1)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Fraction Self Employed") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))

# creating hs and quiv working
fraction_hs_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_hs = sum(high_school_and_equiv == 1) / n()) |> 
  pivot_wider(names_from = daca_eligible, values_from = fraction_hs) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

# plotting hs and equiv
fraction_hs_table |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(0, 0.4)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    labs(title = "Fraction High School and Equivalent") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean()

# creating ged working
fraction_ged_table <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18) |> 
  group_by(survey_year, daca_eligible) |> 
  summarise(fraction_ged = sum(ged == 1) / n()) |> 
  pivot_wider(names_from = daca_eligible, values_from = fraction_ged) |> 
  rename(non_eligible = '0',
         eligible = '1') |> 
  mutate(eligible_minus_non = eligible - non_eligible) |> 
  select(survey_year, eligible_minus_non)

get_table(mex_df, ged)

# plotting fraction ged table
get_table(mex_df, ged) |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se,
                      ymax = eligible_minus_non + se)) +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(-0.1, 0.1)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    geom_hline(yintercept = 0, color = "red") +
    labs(title = "Fraction GED") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))

get_table(mex_df, attended_college)

# plotting fraction college table
get_table(mex_df, attended_college) |> 
  filter(survey_year >= 2008) |> 
  ggplot(aes(x = survey_year, y = eligible_minus_non)) +
    geom_line(size = 1) +
    geom_point() +
    geom_errorbar(aes(ymin = eligible_minus_non - se,
                 ymax = eligible_minus_non + se)) +
    geom_hline(yintercept = 0, color = "red") +
    scale_x_continuous(n.breaks = 20) +
    scale_y_continuous(n.breaks = 10, limits = c(0, 0.4)) +
    geom_rect(xmin = 2012, xmax = 2013, ymin = -Inf, ymax = Inf,
              fill = "gray40", alpha = 0.05) +
    labs(title = "Fraction College") +
    xlab("Year") +
    ylab("Difference Between Eligible and Ineligible") +
    ggthemes::theme_clean() +
    theme(axis.text.y = element_text(size = 12),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_text(size = 18),
          axis.title.x = element_text(size = 18))
```

# Extra

```{r}
mex_df <- mex_df |> 
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18,
         survey_year >= 2008) |> 
  mutate(daca_eligible_label = ifelse(daca_eligible == 1, "daca eligible", "ineligible"))

mex_df |>
  group_by(survey_year, daca_eligible_label) |> 
  summarise(avg_y = mean(pernp)) |> 
  ggplot(aes(x = survey_year, y = avg_y, color = daca_eligible_label)) +
  geom_line() +
  geom_vline(xintercept = 2013) +
  scale_x_continuous(n.breaks = 20) +
  ylim(0, 30000)
```


```{r}
### Creating data summary df ###
mex_df_summary <- mex_df |> 
  mutate(daca_eligible_label = ifelse(daca_eligible == 1, "Daca Eligible", "Ineligible")) |>
  filter(between(birth_year, 1981, 2005),
         qtrbir >= 2,
         age >= 18,
         survey_year >= 2008) |> 
  select(daca_eligible_label, age, age_of_entry, yoep, pernp,
         wkhp, years_of_education, years_living_in_us) |> 
  rename(
    'age of entry' = age_of_entry,
    'year of entry' = yoep,
    income = pernp,
    'hours worked per week' = wkhp,
    'years of education' = years_of_education,
    'years living in US' = years_living_in_us
  )

# creating data summary table
datasummary_balance(
  ~daca_eligible_label,
  data = mex_df_summary,
  title = "Non-citizen Mexicans Ages 18-40: DACA Eligible vs. Ineligible in California",
  stars = c('*' = .1, '**' = .05, '***' = 0.01),
  fmt = 2,
  notes = "***Significant at the 1% level. **Significant at the 5% level. *Significant at the 10% Level.",
  dinm_statistic = "p.value"
)
```

# Daca eligible vs ineligible distribution across time
```{r}
mex_df |> 
  group_by(daca_eligible, daca_eligible_label, survey_year) |> 
  count() |> 
  ggplot(aes(x = survey_year, y = n, fill = daca_eligible_label)) +
  geom_col(position = "dodge") +
  scale_x_continuous(n.breaks = 19)
```

